Look, we can do something

### 2024-11-11 paper.tex
Forgot a frame
Add a reference
Add a comment to pmin
Remove 'improved'

\section{A new pore shape parameter}
We propose a new parameter for shape description which is based on the idea of
numerical assession of bottlenecks in pores. For simplicity, we will assume
that our pore is 1-connected. Suppose you have a round pore in the form close to
a disk and another pore in the form of a dumbbell. We want to develop a
parameter which would be close to 1 for the first pore and close to 0 for the
second. For this purpose we want to visit all points which lie on the pore's
boundary and find a pair of points which form a bottleneck, i.e have a small
distance between each other. Obviously, not all pairs of points which are close
to each other form a bottleneck. Namely, on \cref{fig:parameter} points $A$ and
$B$ form a bottleneck and points $A$ and $C$ do not. If you choose a point $O$
which lies outside of the pore and travel along the pore's boundary in one
direction (either clockwise or counter-clockwise), when you arrive at $A$ you
will see the point $O$ to your left and when you are in $B$ you will see $O$ to
your right (or vice-versa). But if you are at $A$ or $C$ you will see $O$ to the
same side from you. Instead of doing this check for ``bottleneckness'', we add a
penalty coefficient to the distance between points, namely for a pair of points
$(A, B)$ we calculate the following:
\begin{equation}
  p_{AB} = \rho_{AB} / L_{AB}
  \label{eq:intro}
\end{equation}
where $\rho_{AB}$ is the Euclidean distance between $A$ and $B$ and $L_{AB}$ is
the length of a pore's boundary segment with end points $A$ and $B$ (there are
two such segments, you need to take the segment with minimal length). Suppose
$B$ is a set of all points on the pore's boundary. We define a new pore shape
parameter $p$ using \ref{eq:intro}:
\begin{equation}
  p = \inf \{ p_{AB} \ \forall (A, B) \in B^2 : 2 L_{AB} \ge cP \}
  \label{eq:pre-awesomeness}
\end{equation}
The condition on a pair of points $(A, B)$ is that the ``boundary distance''
between them must be no less than $\frac{cP}{2}$ where $P$ is a perimeter of a
pore and $c \in [0,1]$ is some parameter. This parameter is introduced to
prevent small features of the boundary from affecting $p$. In this paper we will
use $c = 0.1$. A value $c = 0$ means there are no restrictions on $(A, B)$.

@@

Although a guaranteed upper bound for $p$ is 1, it's reachable only for
degenerate shapes in the form of line segments. The biggest value for
non-degenerate shapes is reached for a disk and equals to $2/\pi$. Similarly, an
ellipse defined as
\begin{equation*}
  \frac{x^2}{a^2} + \frac{y^2}{b^2} = 1
\end{equation*}
has the biggest known value of $p$ from all shapes with elongation $E = a / b$
(suppose $a \le b$). Using Ramanujan's formula for the perimeter of an ellipse
we get
\begin{equation*}
  p \le p_{ellipse} = \frac{2a}{P/2} \approx \frac{4E}{\pi[3(1+E) - \sqrt{(3E+1)(E+3)}]}
\end{equation*}
where $P$ is a perimeter of the ellipse. By dividing $p$ by $p_{ellipse}$ we
obtain an improved pore shape parameter $P$:
\begin{equation*}
  P = \frac{\pi p [3(1+E) - \sqrt{(3E+1)(E+3)}]}{4E}
\end{equation*}

@@

Suppose the boundary is defined by a parameteric equation
\begin{equation*}
  \left\{
  \begin{aligned}
    x &= f(t) \\
    y &= g(t)
  \end{aligned}
  \right.
\end{equation*}
or, using a shorter notation,
$B = \begin{pmatrix} x_B \\ y_B \end{pmatrix} = \begin{pmatrix} f \\ g \end{pmatrix}(t)$
Suppose, you have two points: $B_1 = \begin{pmatrix} f \\ g \end{pmatrix}(t_1)$
and $B_2 = \begin{pmatrix} f \\ g \end{pmatrix}(t_2)$. The sorting in the step 2
implies that if $t_1 < t_2$, the point $B_1$ must appear before $B_2$ in
\texttt{\%bs}.

@@

Finaly, the function \texttt{pmin} is described in \cref{alg:pmin}. This
algorithms assumes that there are enough boundary points in an image. If, for
example, an image is taken with bad resolution (i.e. pores consist only of a few
pixels) some quality improving techniques must be applied first. It also
exploits the commutative property $p_{AB} = p_{BA}$.
\begin{algorithm}[H]
  \caption{The last step of the algorithm for awesomeness: the function \texttt{pmin}.}
  \label{alg:pmin}
  \begin{algorithmic}[1]
    \Procedure{\texttt{pairwise\_distances}}{\texttt{points}}
    \State \textbf{return} \texttt{map(dist, points, tail(points))}
    \Comment Return the Euclidean distance between $i$-th and $i+1$-th
    points. The function \texttt{dist} calculates the distance between two
    points.
    \EndProcedure
    \Procedure{\texttt{perimeter}}{\texttt{points, c}}
    \State \texttt{p1 $\gets$ foldl1((+), pairwise\_distances(points))}
    \State \texttt{p2 $\gets$ dist(head(points), last(points))}
    \State \textbf{return} \texttt{p1 + p2}
    \EndProcedure
    \Procedure{\texttt{minimize\_on\_head}}{\texttt{acc, p, bds, points}}
    \Comment \texttt{acc} is the best (minimal) parameter found so far,
    \texttt{p} is a perimeter of the boundary, \texttt{bds} is a list of
    distances along the boundary from the head of \texttt{points} to all points
    in \texttt{tail(points)}.
    \Procedure{\texttt{go}}{\texttt{acc, x}}
    \State \texttt{(point, bd) $\gets$ x}
    \State \texttt{\%bd $\gets$ min(bd, p - bd)}
    \If{\texttt{\%bd < c/2 $\cdot$ p}}
    \State \textbf{return} \texttt{acc}
    \Else
    \State \texttt{param $\gets$ dist(point, head(points)) / \%bd}
    \State \textbf{return} \texttt{min(acc, param)}
    \EndIf
    \EndProcedure
    \State \texttt{tmp $\gets$ zip(tail(points), ds)}
    \State \textbf{return} \texttt{foldl(go, acc, tmp)}
    \EndProcedure
    \Procedure{\texttt{pmin}}{\texttt{points}}
    \State \texttt{p $\gets$ perimeter(points)}
    \State \texttt{bds $\gets$ scanl1((+), pairwise\_distances(points))}
    \Procedure{\texttt{go}}{\texttt{acc, points}}
    \State \texttt{(param, bds) $\gets$ acc}
    \State \texttt{\%param $\gets$ minimize\_on\_head(param, p, bds, points)}
    \State \texttt{\%bds  $\gets$ map(subtract(head(bds)), tail(bds))}
    \State \textbf{return} \texttt{(\%param, \%bds)}
    \EndProcedure
    \State \textbf{return} \texttt{foldl(go, (+Inf, bds), tails(points))}
    \Comment For a list \texttt{a = [1,2,3]} \texttt{tails(a)} returns
    \texttt{[[1,2,3], [2,3], [3]]}.
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

The parameter \cref{eq:awesomeness} is scale- and rotation-invariant. The
algorithm presented in this paper changes the parameter by about 10\% when
rotating an image of a pore or scaling it by a factor $0.5-2$. This inaccuracy
is OK for assessing quality of pore clusterization, as can be seen in the next
section.
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{images/rotation-resize.png}
  \caption[]{\highlight{Awesomeness} parameter for the same image under rotation
  (red X marks) and resize (blue dot marks) transforms.}
  \label{fig:rotation-resize}
\end{figure}

@@

The first flaw can be dealt with by replacing the Euclidean distance $\rho_{AB}$
in \cref{eq:intro} with the length $d_{AB}$ of the shortest path between points
which does not cross the pore's boundary (see \cref{fig:improved}). Although the
resulting algorithm will have the same $O(n^2)$ complexity, doing so can
increase computational time significantly. We call the parameter $P'$ with
$d_{AB}$ in the place of $\rho_{AB}$ ``improved awesomeness''. The following
relation always holds: $0 < P \le P' \le 1$.
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{images/improved.png}
  \caption[]{Pairs of points responsible for \highlight{awesomeness} (red) and
    improved \highlight{awesomeness} (green).}
  \label{fig:improved}
\end{figure}

>>>>

\section{A new pore shape parameter}
We propose a new parameter for shape description which is based on the idea of
numerical assession of bottlenecks in pores. For simplicity, we will assume
that our pore is 1-connected. Suppose you have a round pore in the form close to
a disk and another pore in the form of a dumbbell. We want to develop a
parameter which would be close to 1 for the first pore and close to 0 for the
second. For this purpose we want to visit all points which lie on the pore's
boundary and find a pair of points which form a bottleneck, i.e have a small
distance between each other. Obviously, not all pairs of points which are close
to each other form a bottleneck. Namely, on \cref{fig:parameter} points $A$ and
$B$ form a bottleneck and points $A$ and $C$ do not. If you choose a point $O$
which lies outside of the pore and travel along the pore's boundary in one
direction (either clockwise or counter-clockwise), when you arrive at $A$ you
will see the point $O$ to your left and when you are in $B$ you will see $O$ to
your right (or vice-versa). But if you are at $A$ or $C$ you will see $O$ to the
same side from you. Instead of doing this check for ``bottleneckness'', we add a
penalty coefficient to the distance between points, namely for a pair of points
$(A, B)$ we calculate the following:
\begin{equation}
  p_{AB} = \rho_{AB} / L_{AB}
  \label{eq:intro}
\end{equation}
where $\rho_{AB}$ is the Euclidean distance between $A$ and $B$ and $L_{AB}$ is
the length of a pore's boundary segment with end points $A$ and $B$ (there are
two such segments, you need to take the segment with minimal length). Suppose
$B$ is a set of all points on the pore's boundary. We define a preliminary pore
shape parameter $p$ using \ref{eq:intro}:
\begin{equation}
  p = \inf \{ p_{AB} \ \forall (A, B) \in B^2 : 2 L_{AB} \ge cP \}
  \label{eq:pre-awesomeness}
\end{equation}
The condition on a pair of points $(A, B)$ is that the ``boundary distance''
between them must be no less than $\frac{cP}{2}$ where $P$ is a perimeter of a
pore and $c \in [0,1]$ is some parameter. This parameter is introduced to
prevent small features of the boundary from affecting $p$. In this paper we will
use $c = 0.1$. A value $c = 0$ means there are no restrictions on $(A, B)$.

@@

Although a guaranteed upper bound for $p$ is 1, it's reachable only for
degenerate shapes in the form of line segments. The biggest value for
non-degenerate shapes is reached for a disk and equals to $2/\pi$. Similarly, an
ellipse defined as
\begin{equation*}
  \frac{x^2}{a^2} + \frac{y^2}{b^2} = 1
\end{equation*}
has the biggest known value of $p$ from all shapes with elongation $E = a / b$
(suppose $a \le b$). Using Ramanujan's formula for the perimeter of an ellipse
we get
\begin{equation*}
  p \le p_{ellipse} = \frac{2a}{P/2} \approx \frac{4E}{\pi[3(1+E) - \sqrt{(3E+1)(E+3)}]}
\end{equation*}
where $P$ is a perimeter of the ellipse. By dividing $p$ by $p_{ellipse}$ we
obtain a parameter $P$:
\begin{equation*}
  P = \frac{\pi p [3(1+E) - \sqrt{(3E+1)(E+3)}]}{4E}
\end{equation*}

@@

Suppose the boundary is defined by a parameteric equation
\begin{equation*}
  \left\{
  \begin{aligned}
    x &= f(t) \\
    y &= g(t)
  \end{aligned}
  \right.
\end{equation*}
or, using a shorter notation,
$B = \begin{pmatrix} x_B \\ y_B \end{pmatrix} = \begin{pmatrix} f \\ g \end{pmatrix}(t)$
Suppose, you have two points: $B_1 = \begin{pmatrix} f \\ g \end{pmatrix}(t_1)$
and $B_2 = \begin{pmatrix} f \\ g \end{pmatrix}(t_2)$. The sorting in the step 2
implies that if $t_1 < t_2$, the point $B_1$ must appear before $B_2$ in
\texttt{\%bs} (denote it as $B_1 < B_2$).

@@

Finaly, the function \texttt{pmin} is described in \cref{alg:pmin}. This
algorithms assumes that there are enough boundary points in an image. If, for
example, an image is taken with bad resolution (i.e. pores consist only of a few
pixels) some quality improving techniques must be applied first. It also
exploits the commutative property $p_{AB} = p_{BA}$.
\begin{algorithm}[H]
  \caption{The last step of the algorithm for awesomeness: the function \texttt{pmin}.}
  \label{alg:pmin}
  \begin{algorithmic}[1]
    \Procedure{\texttt{pairwise\_distances}}{\texttt{points}}
    \State \textbf{return} \texttt{map(dist, points, tail(points))}
    \Comment Return the Euclidean distance between $i$-th and $i+1$-th
    points. The function \texttt{dist} calculates the distance between two
    points.
    \EndProcedure
    \Procedure{\texttt{perimeter}}{\texttt{points, c}}
    \State \texttt{p1 $\gets$ foldl1((+), pairwise\_distances(points))}
    \State \texttt{p2 $\gets$ dist(head(points), last(points))}
    \State \textbf{return} \texttt{p1 + p2}
    \EndProcedure
    \Procedure{\texttt{minimize\_on\_head}}{\texttt{acc, p, bds, points}}
    \Comment \texttt{acc} is the best (minimal) parameter found so far,
    \texttt{p} is a perimeter of the boundary, \texttt{bds} is a list of
    distances along the boundary from the head of \texttt{points} to all points
    in \texttt{tail(points)}.
    \Procedure{\texttt{go}}{\texttt{acc, x}}
    \State \texttt{(point, bd) $\gets$ x}
    \State \texttt{\%bd $\gets$ min(bd, p - bd)}
    \If{\texttt{\%bd < c/2 $\cdot$ p}}
    \State \textbf{return} \texttt{acc}
    \Else
    \State \texttt{param $\gets$ dist(point, head(points)) / \%bd}
    \State \textbf{return} \texttt{min(acc, param)}
    \EndIf
    \EndProcedure
    \State \texttt{tmp $\gets$ zip(tail(points), ds)}
    \State \textbf{return} \texttt{foldl(go, acc, tmp)}
    \EndProcedure
    \Procedure{\texttt{pmin}}{\texttt{points}}
    \State \texttt{p $\gets$ perimeter(points)}
    \State \texttt{bds $\gets$ scanl1((+), pairwise\_distances(points))}
    \Procedure{\texttt{go}}{\texttt{acc, points}}
    \State \texttt{(param, bds) $\gets$ acc}
    \State \texttt{\%param $\gets$ minimize\_on\_head(param, p, bds, points)}
    \State \texttt{\%bds  $\gets$ map(subtract(head(bds)), tail(bds))}
    \Comment $L_{B_2B_3} = L_{B_1B_3} - L_{B_1B_2}$ if $B_1 < B_2 < B_3$.
    \State \textbf{return} \texttt{(\%param, \%bds)}
    \EndProcedure
    \State \textbf{return} \texttt{foldl(go, (+Inf, bds), tails(points))}
    \Comment For a list \texttt{a = [1,2,3]} \texttt{tails(a)} returns
    \texttt{[[1,2,3], [2,3], [3]]}.
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

The parameter \cref{eq:awesomeness} is scale- and rotation-invariant. The
algorithm presented in this paper changes the parameter by about 10\% when
rotating an image of a pore or scaling it by a factor $0.5-2$ (see
\cref{fig:rotation-resize}). This inaccuracy is OK for assessing quality of pore
clusterization, as can be seen in the next section.
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{images/rotation-resize.png}
  \caption[]{\highlight{Awesomeness} parameter for the same image under rotation
  (red X marks) and resize (blue dot marks) transforms.}
  \label{fig:rotation-resize}
\end{figure}

@@

The first flaw can be dealt with by replacing the Euclidean distance $\rho_{AB}$
in \cref{eq:intro} with the length $d_{AB}$ of the shortest path between points
which does not cross the pore's boundary (see \cref{fig:improved}). Although the
resulting algorithm will have the same $O(n^2)$ complexity, doing so can
increase computational time significantly. We call the parameter $P'$ with
$d_{AB}$ in the place of $\rho_{AB}$ ``improved awesomeness''. The following
relation always holds: $0 < P \le P' \le 1$.
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth,frame]{images/improved.png}
  \caption[]{Pairs of points responsible for \highlight{awesomeness} (red) and
    improved \highlight{awesomeness} (green).}
  \label{fig:improved}
\end{figure}
